### 2005-03-14 Documentation/appendix.tex
PH prepared the manual for the release of 2.0.11 and removed part 3 (modules) again since it becomes pertinent only for version 2.3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{appendix}

\chapter{\GLOBES\ installation}
\label{app:installation}
\index{norm}{Installation|(}

The installation of \GLOBES\ is highly automated and there should not be
any problems on a decently up-to-date GNU/Linux system. The installation
has, however, only been tested on a limited number of platforms. The 
current version has been tested for SuSE, Red Hat, Fedora Core and
a 64bit version of SuSE. It should also compile and work on Mac OS. We 
would appreciate to know your experiences with the installation on different platforms.
Please send an e-mail to <\bugs>.

\section*{Prerequisites for the installation of \GLOBES}
\index{norm}{Installation!prerequisites}

Besides the usual things such as a working \verb`libc`, you need to have
\begin{itemize}
\item[gcc] The GNU compiler collection\\
\verb+gcc.gnu.org+
\item[GSL] The GNU Scientific Library\\
\verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any
ANSI C/C++ compiler, but the \verb+globes+ binary uses the 
\verb+argp+ facility of \verb+glibc+ to parse its command line options.
However, on platforms where \verb+argp+ is lacking \GLOBES\ has replacement code, 
thus it should also work there.

All those libraries are also available as rpm's from the various
distributors of GNU/Linux, see their web sites for downloads. Chances are 
that \verb+gcc+ and GSL are already part of your installation.  For building \GLOBES\ 
from source, however, not only working libaries for above packages are 
needed but also the headers,  especially for GSL. Depending on your
installation, eg. on RedHat/Fedora, of GSL this may require to additionally 
install a rpm-package named \verb+gsl-devel+. If GSL has been installed from the 
tar-ball as provided by \verb`gnu.org` no problems should occur. Furthmore you need 
a working \verb+make+ to build and install \GLOBES.

\section*{Installation Instructions}

\GLOBES\ follows the standard GNU installation procedure.  To compile \GLOBES\
you will need gcc.  After unpacking the distribution
the Makefiles can be prepared using the configure command,
\begin{quote}
{\tt
./configure
}
\end{quote}
You can then build the library by typing,
\begin{quote}
{\tt
make
}
\end{quote}
A shared  version of the library will be compiled by
default.

The library can be installed using the command,
\begin{quote}
{\tt
make install
}
\end{quote}
The default install directory prefix is \verb+/usr/local+.  Consult the
"Further Information" section below for instructions on installing the
library in another location or changing other default compilation
options.

The install target also will install a program with name \verb+globes+ to
\verb+$prefix/bin+ and the files in the \verb+data+ directory of the tar-ball
to \verb+$prefix/share/globes+.
\index{norm}{glb files@{\tt glb}-files!installation}

If you are not using \verb+make install+ you will find the static libary
at \verb+source/.lib/libglobes.a+ which you can copy to any destination.
However keep in mind that the linking command will be somewhat different,
\ie\ you have to specify all the dynamically linked objects besides libglobes.

The default install directory prefix is \verb+/usr/local+.  Consult the
"Further Information" section below for instructions on installing the
library in another location or changing other default compilation
options.

Moreover a config-script called \verb+globes-config+ will be installed. This
script displays all information necessary to link any program with GLoBES.
For building static libaries and linking against them see the corresponding
section of this file.

\section*{Basic Installation}

   The \verb+configure+ shell script attempts to guess correct values for
various system-dependent variables used during compilation.  It uses
these values to create a \verb+Makefile+ in each directory of the package.
Finally, it creates a shell script \verb+config.status+ that
you can run in the future to recreate the current configuration, a file
\verb+config.cache+ that saves the results of its tests to speed up
reconfiguring, and a file \verb+config.log+ containing compiler output
(useful mainly for debugging \verb+configure+).

   If you need to do unusual things to compile the package, please try
to figure out how \verb+configure+ could check whether to do them, and mail
diffs or instructions to \bugs\ so they can
be considered for the next release.  If at some point \verb+config.cache+
contains results you don't want to keep, you may remove or edit it.

   The file \verb+configure.in+ is used to create \verb+configure+ by a program
called \verb+autoconf+.  You only need \verb+configure.in+ 
if you want to change it or regenerate \verb+configure+ using a newer 
version of \verb+autoconf+.

The simplest way to compile this package is:
\begin{enumerate}
\item \verb+cd+ to the directory containing the package's source code and type
     \verb+./configure+ to configure the package for your system.  If you're
     using \verb+csh+ on an old version of System V, you might need to type
     \verb+sh ./configure+ instead to prevent \verb+csh+ from trying to execute
     \verb+configure+ itself.

     Running \verb+configure+ takes awhile.  While running, it prints some
     messages telling which features it is checking for. It also prints
     a reminder for things to do after installation.

\item Type \verb+make+ to compile the package.

\item Type \verb+make install+ to install the programs and any data files and
     documentation.

\item You can remove the program binaries and object files from the
     source code directory by typing \verb+make clean+.  To also remove the
     files that \verb+configure+ created (so you can compile the package for
     a different kind of computer), type \verb+make distclean+.  There is
     also a \verb+make maintainer-clean+ target, but that is intended mainly
     for the package's developers.  If you use it, you may have to get
     all sorts of other programs in order to regenerate files that came
     with the distribution.

\item Since you've installed a library don't forget to run \verb+ldconfig+ !
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Installation without root privilege}
\index{norm}{Installation!w/o root privilege}
\label{inst_noroot}

Install \GLOBES\ to a directory of your choice \verb+GLB_DIR+. This is done by
\begin{quote}
\verb+configure --prefix=GLB_DIR+
\end{quote}
and then follow the usual installation guide. The only remaining problem
is that you have to tell the compiler where to find the header files, and 
the linker where to find the library. Furthermore you have to make sure 
that the shared object files are found during execution. Running 
\verb+configure+
also produces a \verb+Makefile+ in the \verb+examples+ 
subdirectory which can serve as
a template for the compilation and linking process, since all necessary
flags are correctly filled in.

Another solution is to set the environment variable \verb+LD_RUN_PATH+
during linking to \verb+GLB_DIR/lib+ . Best thing is to add this to your
shell dot-file (e.g. \verb+.bashrc+). Then you can use:
A typical compiler command like
\begin{quote}
\verb+gcc -c my_program.c -IGLB_DIR/include/+
\end{quote}
and a typical linker command like
\begin{quote}
\verb#g++ my_program.o -lglobes -LGLB_DIR/lib/ -o my_executable#
\end{quote}
More information on this issue can be obtained by having a look 
into the mentioned  \verb+Makefile+ in \verb+examples+.

\begin{itemize}
\item[CAVEAT] It is in principle possible to have many installations 
on one machine. Especially the situation of having an installation by root and by a user at the same time might occur. However, it is strictly warned against this possibility, since it is \emph{extremely} likely to create some versioning problem at some time!
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{LAPACK \& BLAS requirements}

GLoBES uses one single routine out of LAPACK (\verb^www.netlib.org/lapack/^) named
\verb^zgeev^ to diagonalize the oscillation Hamiltonian in matter. Since LAPACK
is not part of the standard installation of many distributions and we need
only one function, \GLOBES\ comes with a convenience version of the required
routines, which are taken from LAPACK and BLAS (\verb^www.netlib.org/blas^) and have
been translated to C with \verb^f2c^. In case your system has a working installation
of BLAS/LAPACK \GLOBES\ will use it. However, if your system is lacking BLAS/
LAPACK the convenience version will be compiled and linked into \GLOBES. If
for some reason BLAS/LAPACK is installed on the system and one still wishes
to use the convience version use \verb^--enable-lapack-convenience^. Using Fortran
code which has been translated to C with \verb^f2c^ requires a run-time library
called \verb^libf2c^ (on some systems it is called \verb^libg2c^). The configury of 
\GLOBES\ detects the presence of either of the two and takes care of it.
If none of the two is present, a convience library is used. If
for some reason \verb^libf2c^ is installed on the system and one still wishes
to use the convience version use \verb^--enable-libf2c-convenience^.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{GSL requirements}

Sometimes the GNU scientific library is not available or is installed in a non-standard
location. This situation can arise in an installation without root privileges. In this 
case one can specify \verb^--with-gsl-prefix=path_to_gsl^ as option to the \verb^configure^ script.
If one wants to use a shared version of \verb^libgsl^ then one has to make sure that the
linker finds the library at run-time. This can be achieved by setting the environment variable
\verb^LD_LIBRARY_PATH^ correctly, \ie\ (in bash)
\begin{quote} 
\verb^export LD_LIBRARY_PATH=path_to_gsl^
\end{quote}
You also can use a static version of GSL by either building \GLOBES\ with \verb^LDFLAG=-all-static^
or by configuring GSL with \verb^--disable-shared^. In both cases no further actions like setting
any environment variables is necessary.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Distributions}

\subsection*{Fedora Core 3}

FC3 is shipped with a non working version of LAPACK. Thus to
make \GLOBES\ work on FC3, configure \GLOBES\ with \verb'--enable-lapack-convenience', which
replaces the defunct LAPACK with \GLOBES' own version thereof. You can diagnose a 
non-working LAPACK by compiling one of the examples in the \verb^examples^ directory and 
trying to run it. It will hang forever in the function \GLB{SetRates()}.

\subsection*{RedHat -- all versions}

The standard rpm-based installation of GSL does not provide any header files for GSL, 
which are however needed to compile \GLOBES. You have to install an additional rpm-package
called \verb'gsl-devel'. Altnernatively you can install GSL from a tar-ball and use the 
verb'--with-gsl-prefix' option to the configure script of \GLOBES.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Advanced topics}

\subsection*{Building and Using static versions of \GLOBES}

In certain circumstances it may be useful to use a static version of libglobes
or any of the binaries, \eg\ when running on a cluster.

   The \verb+configure+ script accepts the option \verb+--disable-shared+, in which
case only static objects are built, i.e. only a static version of libglobes. 
In case your system does not support shared libraries the \verb+configure+ script 
recognizes this. If you give no options to \verb+configure+, both shared and static 
versions are built and will be installed. All binaries, however, will use dynamic 
linking. If you want to build static binaries, use \verb+LDFLAGS=-all-static+ for building
them.

   Sometimes it is convenient, eg. for debugging purposes, to have a statically
linked version of a program using \GLOBES\, which is easiest achieved by just linking
with \verb+libglobes.a+. If you need a completely statically linked version, please, have
a look at the Makefile in the \verb+examples+ directory.
\begin{quote}
\verb#make example-static#
\end{quote}
produces a statically linked program that should in principle run on most Linuxes.
It should be straightforward to adapt this example to your needs.

All these options rely on \verb+libtool+ and a working \verb`gcc` installation. It seems that \verb`gcc 3.x`
is broken in subtle way which makes it necessary to add a symbolic link in the \verb`gcc` library
directory. The diagnostic for this requirement is that building static programs fails with
the error message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add a symbolic 
link in the same directory where you found it (this requires probably root privileges)
\begin{quote}
\verb#ln -s libgcc.a libgcc_s.a#
\end{quote}

If you can not write to this directory just use the following work around. Add the same link  
as above to the directory where you installed \GLOBES\ into
\begin{quote}
\verb#cd prefix/lib#
\verb#ln -s path_to_libgcc.a/libgcc.a libgcc_s.a#
\end{quote}
and then change back into the \verb+examples+ directory and type
\begin{quote}
\verb#make LDFLAGS=-Lprefix/lib example-static#
\end{quote}
and you are done.

If you don't have libtool on your system either, then you just can use the libtool script in
the \GLOBES\ tar-ball instead. Make sure it is in a place where it is found by \verb+make+.

\subsection*{Compilers and Options}

   Some systems require unusual options for compilation or linking that
the \verb+configure+ script does not know about.  You can give \verb+configure+
initial values for variables by setting them in the environment.  Using
a Bourne-compatible shell, you can do that on the command line like
this
\begin{quote}
\verb+CC=c89 CFLAGS=-O2 LIBS=-lposix ./configure+
\end{quote}
Or on systems that have the `env' program, you can do it like this
\begin{quote}
\verb+env CPPFLAGS=-I/usr/local/include LDFLAGS=-s ./configure+
\end{quote}

\subsection*{Compiling For Multiple Architectures}

   You can compile the package for more than one kind of computer at the
same time, by placing the object files for each architecture in their
own directory.  To do this, you must use a version of \verb+make+ that
supports the \verb+VPATH+ variable, such as GNU \verb+make+.  `cd' to the
directory where you want the object files and executables to go and run
the \verb+configure+ script.  \verb+configure+ automatically checks for the
source code in the directory that \verb+configure+ is in and in `..'.

   If you have to use a \verb+make+ that does not supports the \verb+VPATH+
variable, you have to compile the package for one architecture at a time
in the source code directory.  After you have installed the package for
one architecture, use `make distclean' before reconfiguring for another
architecture.

\subsection*{Installation Names}

By default, \verb+make install+ will install the package's files in
\verb+/usr/local/bin+, \verb+/usr/local/include+, etc.  You can specify an
installation prefix other than \verb+/usr/local+ by giving \verb+configure+ the
option \verb+--prefix=PATH+.

\subsection*{Specifying the System Type}

   There may be some features \verb+configure+ can not figure out
automatically, but needs to determine by the type of host the package
will run on.  Usually \verb+configure+ can figure that out, but if it prints
a message saying it can not guess the host type, give it the
\verb+--host=TYPE+ option.  TYPE can either be a short name for the system
type, such as `sun4', or a canonical name with three fields: 
\verb+CPU-COMPANY-SYSTEM+

See the file \verb+config.sub+ for the possible values of each field.

   If you are building compiler tools for cross-compiling, you can also
use the \verb+--target=TYPE+ option to select the type of system they will
produce code for, and the \verb+--build=TYPE+ option to select the type of
system on which you are compiling the package.

\subsection*{Sharing Defaults}
   If you want to set default values for \verb+configure+ scripts to share,
you can create a site shell script called \verb+config.site+ that gives
default values for variables like \verb+CC+,  \verb+cache_file+, and 
\verb+prefix+.
\verb+configure+ looks for \verb+PREFIX/share/config.site+ if it exists, then
\verb+PREFIX/etc/config.site+ if it exists.  Or, you can set the
\verb+CONFIG_SITE+ environment variable to the location of the site script.
A warning: not all \verb+configure+ scripts look for a site script.

\index{norm}{Installation|)}

{\footnotesize
\input{gpl}

\input{fdl}
}

\end{appendix}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: Manual.tex
%%% TeX-master: "Manual"
%%% End: