### 2007-01-13 Documentation/Manual.tex
JK - Some typos in Manual.tex corrected

\documentclass[a4paper,12pt,twoside]{book}

\sloppy
% \frenchspacing
\setlength{\unitlength}{1cm}
% Do not change following line pdfmanual depends on it!
\input{pdfswitch}
%\usepackage{pdfcolmk}
\usepackage{verbatim}
\usepackage{cite}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{fancyheadings}
\usepackage[footnotesize,bf]{caption}
\usepackage{makeidx}
%Preparations for multiple indices
\usepackage{multind}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ae,aecompl} % only for Type 1 fonts for readable pdf convert

\newcommand{\glbxxx}{hep-ph/0407333}
\newcommand{\glbxxxtwo}{hep-ph/XXYYZZZ}
\newcommand{\capdef}{}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
        \caption[#1]{{ %\itshape
 {\footnotesize #2}}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother

% Index: use "makeindex Manual" to create .ind file
%\makeindex

\makeindex{norm}
\makeindex{constants}
\makeindex{api}
\makeindex{aedl}

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ra}{$\rightarrow$}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\ldm}{\Delta m_{31}^2}
\newcommand{\sdm}{\Delta m_{21}^2}
\newcommand{\deltacp}{\delta_{\mathrm{CP}}}
\newcommand{\stheta}{\sin^2 2 \theta_{13}}

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\newcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\Chapt}{Chapter}
\newcommand{\Chapts}{Chapters}
\newcommand{\Part}{Part}
\newcommand{\App}{Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

\newcommand{\JHFSK}{{\sc JHF-SK}}
\newcommand{\NUMI }{{\sc NuMI}}
\newcommand{\ReactorI}{{\sc Reactor-I}}
\newcommand{\ReactorII}{{\sc Reactor-II}}
\newcommand{\JHFHK}{{\sc JHF-HK}}
\newcommand{\NuFactI}{{\sc NuFact-I}}
\newcommand{\NuFactII}{{\sc NuFact-II}}
\newcommand{\SPL}{{\sc SPL}}
\newcommand{\Beta}{{\sc $\beta$-Beam}}
\newcommand{\NOVA}{{\sc NO$\nu$A}}
\newcommand{\TtoK}{{\sc T2K}}
\newcommand{\TtoHK}{{\sc T2HK}}
\newcommand{\DC}{{Double\sc Chooz}}

\newcommand{\GLOBES}{{\sf GLoBES}}
\newcommand{\AEDL}{{\sf AEDL}}
\newcommand{\EDM}{{\sf EDM}}

\newcommand{\equ}[1]{\eq~(\ref{equ:#1})}
\newcommand{\figu}[1]{\fig~\ref{fig:#1}}
\newcommand{\tabl}[1]{\Tab~\ref{tab:#1}}
\newcommand{\tb}{\hspace*{3ex}}

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLB}[1]{{\tt glb#1}\index{api}{#1@{\tt #1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNS}[1]{\index{api}{#1@{\tt #1}}}

% Set as GLoBES command and add to index:
% Always use GLB when mentioning a command in text!
\newcommand{\GLBC}[1]{{\tt #1}\index{constants}{\tt{#1}}}
% Always use GLBNS when adding a command to the index without showing it!
\newcommand{\GLBNSC}[1]{\index{constants}{{\tt #1}}}
\newcommand{\gq}[1]{\begin{quote} {\tt #1} \end{quote}}
\newcommand{\go}[1]{$\rightarrow$ Output: \begin{quote} {\sf #1} \end{quote}}
\newcommand{\bugs}{{\tt globes@ph.tum.de}}
\definecolor{light}{gray}{0.93}
\definecolor{heavy}{gray}{0.35}
\setlength{\fboxrule}{0.1cm}
\newcommand{\example}[2]{
\begin{table}[p]
\begin{center}
\fcolorbox{heavy}{light}{
\begin{minipage}[t][\textheight][t]{15.5cm}
{\em \underline{Example:} #1}

\vspace*{0.3cm}

#2

\end{minipage}
}
\end{center}
\end{table}
}

\newtheorem{function}{Function}[chapter]

%-- page parameters -------------------------------------------------

\pagestyle{fancyplain}

% Olddefs:

%\addtolength{\textwidth}{-1truecm}
%\addtolength{\oddsidemargin}{1.0truecm}
%\addtolength{\evensidemargin}{-0.3truecm}

% Owndefs:

%\addtolength{\textwidth}{0.0truecm}
%\addtolength{\textheight}{1.8truecm}
%\addtolength{\topmargin}{-0.6truecm}
%\addtolength{\oddsidemargin}{0.65truecm}
%\addtolength{\evensidemargin}{-1.0truecm}

% Headdefs:

\advance \headheight by 3.0truept       % for 12pt mandatory...
\lhead[\fancyplain{}{\thepage}]{\fancyplain{}{\rightmark}}
\rhead[\fancyplain{}{\leftmark}]{\fancyplain{\thepage}{\thepage}}
\cfoot{}

\renewcommand{\chaptermark}[1]{
% chapter title im Seitenkopf
\markboth{\uppercase{\chaptername}\ \thechapter.\ \ #1}
          {\uppercase{\chaptername}\ \thechapter.\ \ #1}}
% section title im Seitenkopf
\renewcommand{\sectionmark}[1]{\markright{\thesection\ \ #1}}

\begin{document}

\frontmatter
\pagenumbering{Roman}
\thispagestyle{empty}
{
\setlength{\parindent}{0cm}

%\begin{titlepage}

\font\fa=cmssbx14 scaled 1440
\font\fb=cmbx10 scaled 1200
\font\fc=cmb10 scaled 1200
\font\fd=cmr10 scaled 1000

{\setlength{\baselineskip}{1.2cm}}

\vspace*{1.5cm}

\begin{center}
{ \Large \bf
{\Huge \GLOBES} \\
General Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and experiment definition manual }

%\vspace*{1.5cm}
%{\large \bf !!!DRAFT IN PROGRESS!!! }
\end{center}

\vspace{0.5cm}

\begin{center}
{\large Patrick Huber$^a$, Joachim Kopp$^b$, Manfred Lindner$^b$, \\ Mark Rolinec$^c$, Walter Winter$^d$}
\end{center}

\vspace{0.3cm}

\begin{center}

\end{center}

\vspace{0.5cm}

\begin{center}
\colorbox{black}{\includegraphics[width=7cm]{.save/earthint}}

\vspace*{-1cm}

{\Huge \textcolor{yellow}{GLoBES}}
\end{center}

\vspace{1cm}

\begin{center}
Version from \today\ for \GLOBES\ 3.0
\end{center}

\vspace*{0.5cm}
\vfill
\hrule

\vspace*{0.1cm}

{\em\small $^{\mathrm{a}}$%
       University of Wisconsin,  Physics Department,
       1150 University Av., Madison, WI 53706, USA}

{\em\small $^{\mathrm{b}}$%
       Max--Planck--Institut f\"ur Kernphysik, 
       Postfach 10~39~80, D--69029 Heidelberg, Germany}

{\em\small $^{\mathrm{c}}$%
       Technische Universit\"at M\"unchen,
       Institut f\"ur Theoretische Physik, Physik--Department,\\
       James--Franck--Strasse, D--85748 Garching, Germany}

{\em\small $^{\mathrm{d}}$%
       Universit\"at W\"urzburg, 
       Lehrstuhl f\"ur theoretische Physik II, \\
       Institut f\"ur theoretische Physik und Astrophysik, 
       Am Hubland,
       D-97074 W\"urzburg, Germany}

%\end{titlepage}

%\title{}
%\date{}
%\author{}
%\maketitle

}

\clearpage
\thispagestyle{empty}
\bigskip
\begin{quote}
    Copyright \copyright  2004--2007  The GLoBES Team.
    Permission is granted to copy, distribute and/or modify this document
    under the terms of the GNU Free Documentation License, Version 1.2
    or any later version published by the Free Software Foundation;
    with the invariant Sections ``Terms of usage of \GLOBES" 
    and ``Acknowledgments'', no Front-Cover Texts, and no Back-Cover Texts.
    A copy of the license is included in the section entitled "GNU
    Free Documentation License".
\end{quote}
\bigskip

\cleardoublepage
\setcounter{page}{1}

\chapter*{What is \GLOBES ?}

\vspace*{-4ex}
\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline and reactor experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language (AEDL), which allows to describe most classes 
of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for GNU/Linux. Since the source code is included,
the port to other operating systems is in principle possible. The software
as well as up-to-date versions of this manual can be found at this URL:
 \verb+http://www.mpi-hd.mpg.de/~globes+

\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
Such experiments are neutrino beam experiments and reactor experiments. 
Geometrical effects of a source distribution, such as in the sun or the 
atmosphere, can not be described. In addition, sources with a physically 
significant time dependence, such as supernov\ae, can not be studied. It 
is, however, possible to simulate beams with bunch structure, since the 
time dependence of the neutrino source is physically only important to suppress backgrounds.

On the experiment definition side, either built-in neutrino fluxes
(\eg, neutrino factory, $\beta$-Beam) or arbitrary fluxes can be used. Similarly,
arbitrary cross sections, energy dependent efficiencies, the
energy resolution function, the considered oscillation channels, 
backgrounds, and many other features can be specified. 
For the systematics, energy
normalization and calibration errors can be simulated in a straightforward way,
or the systematics can be completely user-defined (Version 3.0 and higher). Note that
the energy ranges and windows, as well as the bin widths can be
(almost) arbitrarily chosen, which means that variable bin
widths are allowed. Together with \GLOBES\ comes a number of
pre-defined experiments in order to demonstrate the capabilities
of \GLOBES\ and to provide prototypes for new experiments.

With the C-library, one can extract the $\Delta \chi^2$ for all defined 
oscillation channels for an experiment or any combination of experiments.
Of course, also low-level information, such as oscillation
probabilities or event rates, can be obtained. \GLOBES\ includes the
simulation of neutrino oscillations in matter with arbitrary matter 
density profiles, as well as it allows to simulate the matter density
uncertainty. As one of the most
advanced features of \GLOBES , it provides the technology to 
project the $\Delta \chi^2$, which is a function of all oscillation
parameters, onto any subspace of parameters by local minimization. 
This approach allows the inclusion of multi-parameter-correlations,
where external input (\eg, from solar experiments) can be imposed, too.
Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, all oscillation parameters can be kept free to precisely localize 
degenerate solutions.

\chapter*{Terms of usage of \GLOBES}

\subsection*{Referencing the \GLOBES\ software}

\GLOBES\ is developed for academic use. Thus, the \GLOBES\ Team would
appreciate being given academic credit for it. Whenever you use \GLOBES\
to produce a publication or a talk indicate that you have used \GLOBES\ and
please cite the following references~\cite{globes_paper,globes_paper_two}
\begin{quote}
P. Huber, M. Lindner and W. Winter\\
Simulation of long baseline neutrino oscillation experiments with \GLOBES\\
Comput. Phys. Commun. 167 (2005) 195, arXiv:\glbxxx,
\end{quote}

\begin{quote}
P. Huber, J. Kopp, M. Lindner, M. Rolinec, and W. Winter\\
TO BE DEFINED, NEEDS TO BE UPDATED [WW]\\
arXiv:\glbxxxtwo,
\end{quote}
but \emph{not} this manual. This manual itself is not a scientific 
publication and will not be submitted to a scientific journal. 
It will evolve during time since it is intended for 
regular revision. Besides that, many of the data which are used by \GLOBES\ 
and distributed together with it should be properly referenced. 
For details see below.

Apart from that, \GLOBES\ is free software and open source, \ie, it is 
licensed under the GNU Public License.

\subsection*{Referencing the data in \GLOBES}
\label{ref_data}

\index{norm}{Referencing!data in \GLOBES}
\GLOBES\ wouldn't be useful without having high quality input data.
Much of these input data have been published elsewhere and the authors
of those publications would appreciate to be cited whenever their work
is used. It is solely the user's responsibility 
to make sure that he understands where the input material for \GLOBES\ comes
from and if additional work has to be cited in addition to the 
\GLOBES\ papers~\cite{globes_paper,globes_paper_two}. To assist with this task, we provide  the necessary information for the data coming along together with \GLOBES.

When using the built-in Earth matter density profile, the 
original source is \Ref~\cite{Stacey}.

All files ending with \verb+.dat+ or \verb+.glb+ 
in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have on top a comment field which clearly indicates which works should be
cited when using a certain file. Make sure that dependencies are correctly
tracked, \ie , in some cases files included by other files need to be 
checked, too (for example, cross section or flux files). One can use 
the \verb+-v3+ option to \verb+globes+ to see which files
are included.

It is recommended that you use the same style for your own input files, since then, in case they are distributed, everybody will know how to correctly reference your work.

\chapter*{What is new in \GLOBES ?}

Here we briefly summarize the main changes of the new GLoBES version. For details, please
refer to the respective parts of the manual. Please note that any new \GLOBES\ version is
compatible with older versions, \ie, old application software and \AEDL\ files should, with minor
modifications, run with the new version as well.  However, some function names and features will
evolve during time, which means that outdated features may not be documented anymore.

\section*{Version 3.0}
\index{norm}{Version 3.0}

Here comes a summary of the most important changes in this version for users of earlier
versions of \GLOBES .

\subsection*{New features}

\begin{itemize}
 \item
  User-defined systematics, which can be used to simulate reactor experiments \etc ; see \Sec~\ref{sec:userchi}
and~\ref{sec:rules}
 \item
  User-defined priors to include arbitrary external information in the $\chi^2$ before
marginalization over the oscillation parameters; see \Sec~\ref{sec:userdefined} 
\item
  Non-standard physics support; see \Chapt~\ref{chapt:nsphysics}
\item
  Beta beam fluxes available as built-in fluxes; see \Sec~\ref{sec:source}
\item
  Enhanced support for parallelization, such as Condor; see, \eg, page~\pageref{sec:condor}
\item
  Updated AEDL files; see \Tab~\ref{tab:experiments}
\item
 New \AEDL\ features, such as the support of lists as variables and an interpolation
routine; see \Sec~\ref{sec:advaedl}
\item
 Clean-up of inconsistencies, such as an overall (internal) normalization factor in the flux files;
see, \eg, page~\pageref{app:flux}
\item
  Faster probability engine, easier installation (internal changes)
\item
  Experimental feature: Alternative minimizer provided, which is usually faster than
the standard minimizer; see \Chapt~\ref{chapt:experimental}
\end{itemize}

\subsection*{Major changes}

Most of the modifications should not require that old pieces of software be changed. 
However, the following changes are relevant:
\begin{itemize}
\item
 AEDL requires now that {\tt \$version} be used to define the minimum version number this AEDL
file is to be used with. With this requirement one can easily avoid that new AEDL
files with new features be used with old versions of \GLOBES\ which may not recognize these features.
\item
 Some of the earlier AEDL files have been updated, changed names, or have been removed. In addition, new files have been added. Although old AEDL files will run as usual for compatibility, they will not be
supported by the \GLOBES\ team anymore. You should make sure to keep these files when updating \GLOBES . 
\end{itemize}

\subsection*{Minor changes}

Here we document the most important changes which should not affect older software:
\begin{itemize}
\item
 Functions and constants renamed for consistency:
\begin{itemize}
\item
 {\tt glbChiTheta} $\rightarrow$ {\tt glbChiTheta13}
\item
 {\tt glbChiThetaDelta} $\rightarrow$ {\tt glbChiTheta13Delta}
\item
 {\tt glbChiDms} $\rightarrow$ {\tt glbChiDm21}
\item
 {\tt glbChiDm} $\rightarrow$ {\tt glbChiDm31}
\item
 {\tt GLB\_DM\_SOL} $\rightarrow$ {\tt GLB\_DM\_21}
\item
 {\tt GLB\_DM\_ATM} $\rightarrow$ {\tt GLB\_DM\_31}
\item
 {\tt glbSetStartingValues} $\rightarrow$ {\tt glbSetCentralValues}
\item
 {\tt glbGetStartingValues} $\rightarrow$ {\tt glbGetCentralValues}
\item
 {\tt glbGetProfileData} $\rightarrow$ {\tt glbGetProfileDataInExperiment}
\end{itemize}
\item
 Rate access changed; see \Sec~\ref{sec:event_rates}
\item
 Systematics concept changed, concept of error dimensions removed; {\tt @backgroundcenter} removed, central values for all systematics parameters
now zero; see \Sec~\ref{sec:rules}
\end{itemize}

\cleardoublepage
\tableofcontents

\cleardoublepage
\setcounter{page}{1}
\pagenumbering{arabic}

\chapter*{How to use this manual}
\addcontentsline{toc}{chapter}{How to use this manual}

As it is illustrated in \figu{GLOBES}, \GLOBES\ consists 
of several modules.

\begin{figure}[bht]
\begin{center}
\includegraphics[width=16cm]{GLOBES}
\end{center}
\caption{\label{fig:GLOBES} Different modules in \GLOBES .}
\end{figure}
%
\AEDL (``Abstract Experiment Definition Language'') is a language
to define experiments in the form of ordinary text files. One or more of 
the resulting \AEDL\ files can then be processed together with supporting 
flux or cross section files by the user interface. The user interface
is a C-library, which loads one or more \AEDL\ file(s)
containing the experiment definition(s). The user interface is linked 
against the application software, and provides the user interface functions
for the intended experiment simulation.

The application 
software is, except for some example files, not part of \GLOBES , since
the evaluation of the experiment performance is often a matter of taste
and definition. In addition, the algorithms depend, especially for
high-precision instruments, very much on the oscillation parameters.
In general, it is quite simple to simulate superbeams and reactor
experiments. However, because of the more complicated topology, the
simulation of neutrino factories is much more difficult. In order
to demonstrate some of these difficulties, we present in this manual mainly
examples with neutrino factories. These examples can be found in
\Part~\ref{part:1} within the boxed pages. As complete files, they
are also available in the \GLOBES\ software package.

The \GLOBES\ software may have two target groups: 
Physicists, who are mainly interested in optimizing the potential
of specific experimental setups, and others, who are mainly
interested in the physics potential of different experiment types
from a theoretical point of view.
For the first group, \AEDL\ could be the most interesting aspect of
\GLOBES , where the user interface is only a tool to obtain specific
parameter sensitivities. In this case, \GLOBES\ could serve as a
unified tool for the comparison and optimization of different experiment
 setups on equal footing, where
it is the primary objective to simulate the experiments as accurate
as possible. In addition, changes in experimental parameters, such as
efficiencies or energy resolutions, can quickly be tested.
%
For the second user group, the pre-defined 
experiment definition files might already be sufficient to test
new conceptual approaches, and the user interface is the most interesting
aspect for sophisticated applications including correlations,
degeneracies, and multi-experiment setups. In either case, the \GLOBES\
software could serve as a platform for the exchange of experiment
definitions, and for an efficient splitting of work between
experimentalists and theorists.

The user interface functions are described in \Part~\ref{part:1} of 
this manual, which is the ``user's manual''. In there, first of all a 
short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
have an overview over \GLOBES .
After that, the user
interface is successively introduced from very basic to more sophisticated
functions. Eventually, it is demonstrated how one can change many
experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
information. We recommend that everybody interested in \GLOBES\ should
become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
 and some of the examples on the boxed pages. The examples can be 
 directly compiled 
 from the respective directory in the \GLOBES\ software package.

In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
introductory chapter, all functions are defined in greater detail.
This part might be more interesting for the experimental users who
want to modify or create \AEDL\ files. A useful tool in this context
is the executable program \verb+globes+, which returns event rates and other
information for individual \AEDL\ files without further programming. 
For example, flux normalizations can with this tool be easily adjusted 
to reproduce the event rates of a specific experiment. It is described
in the last chapter of \Part~\ref{part:2}.

% THE FOLLOWING HAS CHANGED: C99, COMPILED WITH GCC; MAYBE ONLY PROBLEMS WITH C89
%
% \paragraph{Note:} All examples for application software in C do require
% a C++ compiler to be properly compiled. For pedagogical reasons, variable
% declarations are done at that place where the variable is needed for the
% first time, which is at variance with C syntax but not with C++ syntax.
% That is the only way in which the examples deviate form ISO C. Moreover
% the actual numerical values of the results of the examples may be different
% from the ones in this manual.
%
% Comment by JK: A pedantic C89 compile will reject the example files (it would
% also reject GLoBES itself). With its default settings, gcc is relatively
% tolerant, so everything works with gcc. I assume that also the Intel
% compiler would be OK, but maybe Borland, Microsoft, or other compilers would
% produce problems.

\mainmatter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART I: User's manual
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{part1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART II: Experiment definition module
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{part2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\backmatter
\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

We would like to thank Martin Freund, who wrote the very first
version of a three-flavor matter profile treatment many years ago,
and Thomas Schwetz, who has been pushing the software to the edge in the 
past few years. Furthermore, we would like to thank Tommy Ohlsson,
Toshihiko Ota, and Julian Skrotzki for using and testing unpublished new features of the software.
PH is especially thankful for the invaluable advice of Thomas Fischbacher on
many design issues in the early stage of the project.
% In addition, we would like to thank
% Mark Rolinec for his help to translate the experiment descriptions
% into \AEDL, and for creating the illustrations in this manuscript.
Finally, thanks to all the people who have been pushing this project
for many years, to the ones who have been continuing asking for the
publication of the software, and the referees of several of our
papers for suggestions which lead to improvements in the software.

This work and the development of \GLOBES\ have over the years been supported by:
\begin{itemize}
\item Technische Universit\"at M\"unchen [All authors]
\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft [All authors]
\item Studienstiftung des Deutschen Volkes [JK, WW]
\item Institute for Advanced Study, Princeton [WW]
\item W. M. Keck Foundation [WW]
\item National Science Foundation [WW]
\item University of Wisconsin, Madison [PH]
\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
\item Universit\"at W\"urzburg [WW]
\end{itemize}

\input{appendix}

% Bibliography:

\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{./OurBibTeX}
\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Index:

\chapter{Indices}
\printindex{api}{API functions}

\printindex{constants}{API constants \& macros}

\printindex{aedl}{AEDL reference}

\printindex{norm}{Index}

\end{document}