### 2023-09-27 Distribution/examples/Dune_EFT_study2/Fluxes_smeft/ProductionCoefficients_DUNE_separated/extraction.Rmd

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}

vector1 = seq(2,151,6)
vector2 = seq(3,151,6)
vector3 = seq(4,151,6)
vector4 = seq(5,151,6)
vector5 = seq(6,151,6)
vector6 = seq(7,151,6)

# Set the path to the directory where the input files are located
input_dir <- "."

# Get a list of all the txt files in the directory
input_files <- list.files(input_dir, pattern = "\\.txt$", full.names = TRUE)

# Loop through the input files and extract the desired columns
for (file in input_files) {
  # Read the input file
  data <- read.table(file, header = FALSE, stringsAsFactors = FALSE)

  # Extract the desired columns and save them to separate files
  output_file1 <- paste0("plus_", basename(file))
  output_file2 <- paste0("minus_", basename(file))
  vec = c(vector1,vector2,vector3)
  vec2 =  c(vector4,vector5,vector6)
  columns1 <- c(vector1,vector2,vector3)[order(vec, decreasing = FALSE)]
  columns2 <- c(vector4,vector5,vector6)[order(vec2, decreasing = FALSE)]

  data1 <- data[, c(1, columns1)]
  data2 <- data[, c(1, columns2)]

  write.table(data1, file = output_file1, sep = "\t", quote = FALSE, row.names = FALSE,col.names = FALSE)
  write.table(data2, file = output_file2, sep = "\t", quote = FALSE, row.names = FALSE,col.names = FALSE)
}

```

```{r}
data <- read.table("ProdALLDpDecay.txt", header = FALSE)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.