### 2023-09-27 Distribution/examples/Dune_EFT_study2/xsec/detection_coefficients/interpolation.Rmd

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
# Load required libraries

library(ggplot2)

v <- c(0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 0.125, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 10, 10, 10, 10, 10, 10, 10, 10, 10)

energy_values <- cumsum(v)

# List all .txt files in the current directory
file_list <- list.files(pattern = "*.txt")

# Process each file and store the results in a list
result_list <- lapply(file_list, function(file) {

  # Read the data file
  data <- read.table(file, header = FALSE)

  # Interpolate and evaluate each function
  interpolated_data <- data.frame(V1 = energy_values)
  for (i in 2:ncol(data)) {
    f <- approxfun(10^(data$V1), data[[i]], method = "linear")
    interpolated_data[[i]] <- f(energy_values)
  }

  # Replace NaN values with 0
  interpolated_data[is.na(interpolated_data)] <- 0

  # Save the interpolated data to a new file
  write.table(interpolated_data, paste0("modified_",file,sep=""),
              row.names = FALSE, col.names = FALSE, quote = FALSE)

  return(interpolated_data)
})

# Combine all interpolated data for plotting
combined_data <- do.call(rbind, result_list)

# Plot the combined interpolated functions
ggplot(combined_data, aes(x = V1)) +
  geom_line(aes(y = V2), color = "red") +
  geom_line(aes(y = V3), color = "blue") +
  geom_line(aes(y = V4), color = "green") +
  xlab("Energy") +
  ylab("Function Value")

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.